---
title: "AE 03: Duke Forest + data visualization"
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest.
As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: true

duke_forest <- duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"), "Garage", "No garage"))
```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

duke_forest |>
  ggplot(aes(x = price, fill = garage)) +
  geom_histogram(binwidth = 50000, show.legend = FALSE) +
  facet_wrap(~ garage) +
  labs(
    title = "Distribution of House Prices by Garage Availability",
    x = "Price (USD)",
    y = "Count"
  ) +
  theme_minimal()
```

Houses with garages tend to have higher prices overall compared to those without garages, as indicated by the distribution shifting toward higher values. The spread of prices for houses with garages also appears wider, suggesting more variability in higher-priced homes. This suggests that having a garage is associated with higher home prices, although it may also reflect other features correlated with garages.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given marker larger houses will be priced higher.
It's also expected that the age of the house will have an effect on the price.
However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive.
So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area

duke_forest |>
  ggplot(aes(x = area, y = price, color = year_built)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "House Price vs Area Colored by Year Built",
    x = "Area (square feet)",
    y = "Price (USD)",
    color = "Year Built"
  ) +
  theme_minimal()
```

Claim 1: Larger houses are priced higher

There is a clear positive relationship between area and price, with larger houses generally having higher prices. The upward trend in the scatterplot and smooth line supports this claim.

Claim 2: Newer houses are priced higher

Newer houses (indicated by color) tend to cluster toward higher prices, although there is some overlap with older homes. This suggests that newer houses are often more expensive, but the relationship is not perfectly consistent.

Claim 3: Bigger and more expensive houses tend to be newer

Larger and more expensive houses often appear to be newer, as indicated by their color distribution in the upper-right portion of the plot. However, there are exceptions, suggesting that while the trend exists, it is not absolute.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
